---
title: "The `glacier` Console Interface"
author: "Lilin Yang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The `glacier` Console Interface}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
```{r setup}
library(glacier)
```

This is a tutorial on how to use the functions contained in the `glacier`
package in the context of an interactive R session. The use of all exported
functions will be demonstrated, but additional functionality can be explored by
browsing the package documentation.

## Annotations
```{r}
anno_path <- system.file("extdata", "ex_anno.csv", package = "glacier")
```
Annotations must first be imported from annotation files. Annotation files
contain the gene set name, followed by a list of annotations, and finally an
optional description. Below is an example of an annotation file, notice that it
contains a header and a description:

```{r echo = FALSE}
knitr::kable(read.csv(anno_path, header = TRUE), "pipe")
```

Annotations can thus be imported:

```{r}
anno <- import_annotations(anno_path, ",", TRUE, c(2, 4), 5)
```

## Databases
```{r}
data_path <- system.file("extdata", "ex_data.csv", package = "glacier")
msig_path <- system.file("extdata", "ex_msig.xml", package = "glacier")
```

Databases must also be imported from database files. Database files contain the
gene set name, followed by a list of genes, and finally an optional description.
Below is an example of a database file, notice that it contains neither a header
nor a description:

```{r echo = FALSE}
knitr::kable(read.csv(data_path, header = FALSE), "pipe")
```

Databases can thus be imported:

```{r}
data <- import_database(data_path, ",", FALSE)
```

Databases in the Molecular Signatures Database (MSigDB) XML format can also be
imported. These databases will allow filtering of categories and organisms later
on. For example:

```{r}
msig <- import_msigdb(msig_path)
```

## Genes
Genes and gene values can be imported from the console. For example:

```{r}
gene_only <- process_input_text("FCN1 FTL CLU")
```

Gene values can also be placed alongside gene names. However, these values will
only be used when creating the overlap heatmap, and will not be used in the
calculation of overlap statistics. For example:

```{r}
gene_vals <- process_input_text("FCN1 0.1 FTL 0.8 CLU 0.05")
```

Gene values can also be imported from Seurat objects. At least one cluster of
interest must be specified. For example:

```{r}
seurat_path <- system.file("extdata", "ex_seurat.rds", package = "glacier")
seurat_data <- readRDS(seurat_path)

gene_seurat <- process_input_seurat(seurat_data, 0)
```

## Computation
Overlap statistics can then be calculated. For example:

```{r}
stats <- compute(gene_only, anno, data)
```

Upon completion, a list of annotation-gene matches, along with the overlap
statistics will be returned. The following is an example of the statistics
calculated:

```{r echo = FALSE}
knitr::kable(stats$stats, "pipe")
```

## Plot statistics
The results of the overlap calculation can then be plotted:

```{r fig.height = 3, fig.width = 7}
plot_stats(stats$stats, "Fold Enrichment", "Adjusted P-value")
```

## Plot overlap
The list of annotation-gene matches can also be plotted. Here, the gene values
from the input above are used for coloring the boxes:

```{r fig.height = 3, fig.width = 7}
plot_overlap(stats$matches, "Gene Value", gene_vals, stats$stats)
```
